### 2024-11-02 paper/2-background.tex
Minor fix
spectec background version 1

% rigorous standardization -> spec is rigorous: formal notation & prose notation
To mitigate this problem, WebAssembly has been standardized very rigourously by
the W3C~\cite{wasm-w3c}.
Especially, WebAssembly specification is written very rigorously.
It describes the semantics of WebAssembly in two forms: formal notation and
prose notation.
Formal notation uses mathematical rules to compactly describe the semantics,
and it is used for proofs such as type soundness.
On the other hand, prose notation uses psudocodes to explain the semantics step
by step.
Most WebAssembly users and engine developers are not familiar with
mathematical rules, so they utilize the prose notation.

@@

% control structure
The stack stores not only the input and output values of the instructions but also
structures related to the control flow.
A \textit{call frame} and a \textit{label} are used for function call and
branching, respectively.
In exception handling proposal, there is a \textit{handler} for exception
handling, and other strucutres could be introduced in the future.
How these control structures are used for the control flow will be discussed in
\cref{sec:control flow in official prose}. \\
\red{TODO: remove dangling section header}

@@

>>>>

% rigorous standardization -> spec is rigorous: formal notation & prose notation
To mitigate this problem, WebAssembly has been standardized very rigourously by
the W3C~\cite{wasm-w3c}.
Especially, WebAssembly specification document is written very rigorously.
It describes the semantics of WebAssembly in two forms: formal notation and
prose notation.
Formal notation uses mathematical rules to compactly describe the semantics,
and it is used for proofs such as type soundness.
On the other hand, prose notation uses psudocodes to explain the semantics
through a step by step instruction.
Most WebAssembly users and engine developers are not familiar with
mathematical rules, so they utilize the prose notation.

@@

% control structure
The stack stores not only the input and output values of the instructions but also
structures related to the control flow.
A \textit{call frame} and a \textit{label} are used for function call and
branching, respectively.
In exception handling proposal, there is a \textit{handler} for exception
handling, and other strucutres could be introduced in the future.
How these control structures are used for the control flow will be discussed in
\cref{sec:control flow in official prose}.

@@

% SpecTec
SpecTec is a tool for \textit{mechanizing} WebAssembly specification.
Specification mechanization is a technique that treats the specification as
data that can be manipulated by a computer.
SpecTec provides a domain specific language(DSL) for specification, and it
automatically generates a specification documents including the semantics
described in formal notation and prose noatation from the DSL.
The DSL is a declarative language that resembles textbook style notation, and
it is type checked to prevent meta-level errors such as notation misuses and
dimension mismatches.

% prose language
However, The DSL cannot generate the prose notation just as it is because it is
declarative.
Therefore, SpecTec also defines another imperative language for the prose
notation, and translates the DSL into the prose language to generate the prose
notation.

% interpreter
What is especially intriguing here is that the prose language is executable.
As a result, the prose program generated from the DSL serves as a WebAssembly
interpter.
That is to say, if WebAssembly semantics is written in the DSL, SpecTec
automatically generates the WebAssembly interpreter following the written
semantics.
The generated interpreter passes all the execution tests in the official test
suites except for the tests related to the infinte loop.

% other backends
Furthermore, SpecTec automatically translates the DSL into Coq\cite{il2coq} for
mathematical proofs and it is working on generating other artifacts such as
test suites required to the WebAssembly standization.
